\Section{Authors}
%
\begin{description}
\item[Mario A. Storti (CIMEC)]
PETSc-FEM kernel, NS and AdvDif modules. 
\item[Norberto M. Nigro (CIMEC)]
PETSc-FEM kernel, NS and AdvDif modules, multi-phase flow. 
\item[Rodrigo R. Paz (CIMEC)]
AdvDif module, AdvDif module, hydrology module, compressible flow,
fluid-structure interaction.
\item[Lisandro Dalc\'\i{}n (CIMEC)]
PETSc-FEM Kernel, Python extension language project scripting
extension, linear algebra, preconditioners.
\item[Ezequiel López (CIMEC)]
Mesh relocation algorithms (mesh-move). 
\end{description}
%
Also many people from the following institutions has contributed
directly or indirectly to the generation of this code:
%
\begin{description}
\item[CIMEC] \emph{International Center for
Computational Methods in Engineering, Santa Fe, Argentina} 
\url{http://www.cimec.org.ar}
\item[INTEC] \emph{Instituto de Desarrollo Tecnol\'ogico para la
  Industria Qu\'\i{}mica}
\url{http://www.intec.unl.edu.ar}
\item[CONICET] \emph{Consejo Nacional de Investigaciones
  Cient\'\i{}ficas y T\'ecnicas}
\url{http://www.conicet.gov.ar}
\item[UNL] \emph{Universidad Nacional del Litoral}
\url{http://www.unl.edu.ar}
\end{description}

%<*>---<*>---<*>---<*>---<*>---<*>---<*>---<*>---<*>---<*>---<*> 
\Section{Grants received}

The development of this code has been developed under financement from
the following grants:

%% To generate this list run $ make projlist in texstuff/curric with
%% LANG=en. Also edit and strip the section title. 
\input projlist

\Section{Symbols and Acronyms}

\SSection{Acronyms} 
\begin{description}
\item[CFD:] Computational Fluid Dynamics
\item[DX:] IBM Data Explorer
\item[ePerl:] embedded Perl
\item[FDM:] Finite Difference Method
\item[FEM:] Finite Element Method
\item[IISD:] Interface Iterative -- Sub-domain Direct method
\item[KWM:] Kinematic Wave Model (see \ref{sec:hydro})
\item[LES:] Large Eddy Simulation
\item[MPI:] Message Passing Interface
\item[OOP:] Object Oriented Programming
\item[Perl:] Practical Extraction and Report Language
\item[PETSc:] Portable Extensible Toolkit for Scientific computations.
\item[SUPG:] Streamline Upwind/Petrov Galerkin
\item[ANN:] Approximate Nearest Neighbor problem. Also refers to the
library developped by David Mount and Sunil Arya
(\url{http://www.cs.umd.edu/~mount/ANN}). 

\end{description}

\Section{Symbols} 

\begin{itemize}
   \item $u,v=$ Streamwise and normal components of velocity.
\end{itemize}

\newpage
\bibliography{petscfem}

\begin{flushleft}
\printindex
\end{flushleft}
